The interactions of rational, pragmatic agents\\ lead to efficient language structure and use

### 2019-02-02 paper/paper.Rmd
BP final read through, now submitting.

### Speakers and listeners
The Rational Speech Act framework (RSA) is a computational-level theory of pragmatic language use, which has produced good fit to human communication behavior across a range of language phenomena [@FrankGoodman2012a; @GoodmanFrank2016a]. RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a]. We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting (see \href{https://bit.ly/2RBSGcU}{SI}).\par

@@

Piantadosi et al. [-@Piantadosi2011a] argue just the opposite, claiming that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*. Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low. In particular, context (or common ground) can provide useful information for disambiguation.\par

As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), differing in cost, and two languages ($\ell_1$, $\ell_2$), describing different utterance-meaning mappings. In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$). By contrast, in language $\ell_2$, $u_1$ can only refer to $m_1$ and $u_2$ can only refer to $m_2$ ($[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$). While it is cheaper for a speaker to use $\ell_1$ (because speaking is always low-cost), it is more difficult for a listener (because $u_1$ is ambiguous). Crucially, however, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par

@@

We conduct $N=2000$ simulations. For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ ($U$ is our set of utterances and $M$ our set of meanings). Recall that languages are boolean matrices and a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every row has some non-zero assignment). For a given simulation, the goal is to find the language $\ell^*$ which minimizes our objective and then check to see if that language contains ambiguity.\par

@@

We are interested in how each speaker-model uses ambiguity over the discourse. A speaker strategy that is mutually efficient for both agents should avoid ambiguity until sufficient contextual information has accumulated. We should expect this to be reflected in our \textit{Full pragmatics} model who reasons about the listener and discourse history. By contrast, a speaker-optimal model who does not consider the listener should greedily use ambiguous utterances (\textit{No Pragmatics} model), while a listener-optimal model should avoid ambiguity entirely (\textit{Partial pragmatics} model).\par

@@

Figure 3, panel (A) shows the empirical probability a speaker uses an ambiguous utterance as a function of discourse position. The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse. The \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely. But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds. Panel (C) tracks our CE objective for each model over the discourse. Note that the objective decreases for all three models, primarily driven by the listeners updating their beliefs about the actual topic ($P(C=c_{current}|D)$). However, the objective declines more quickly under the \textit{Full} and \textit{Partial pragmatic} speakers as listener agents are better able to infer the correct context. Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs. Panel (B) tracks speaker effort, which remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines. But, effort declines in the \textit{Full pragmatic} model as speakers increasingly rely on ambiguous material later in the discourse.\par

>>>>

### Speakers and listeners
The Rational Speech Act framework (RSA) is a computational-level theory of pragmatic language use, which has produced good fit to human communication behavior across a range of language phenomena [@FrankGoodman2012a; @GoodmanFrank2016a]. RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a]. We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speakers and listeners in the reference game setting (see \href{https://bit.ly/2RBSGcU}{SI}).\par

@@

Piantadosi et al. [-@Piantadosi2011a] argue just the opposite, claiming that ambiguity is an efficient property of any communication system in which *communication is contextualized*. Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low. In particular, context (or common ground) can provide useful information for disambiguation.\par

As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), differing in cost, and two languages ($\ell_1$ and $\ell_2$), describing different utterance-meaning mappings. In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$). By contrast, in language $\ell_2$, $u_1$ can only refer to $m_1$ and $u_2$ can only refer to $m_2$ ($[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$). While it is cheaper for a speaker to use $\ell_1$ (because speaking is always low-cost), it is more difficult for a listener (because $u_1$ is ambiguous). Crucially, however, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par

@@

We conduct $N=2000$ simulations. For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ ($U$ is our set of utterances and $M$ our set of meanings). Recall that languages are boolean matrices and a language $\ell \in L$ is "valid" so long as each possible meaning $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every row has some non-zero assignment). For a given simulation, the goal is to find the language $\ell^*$ which minimizes our objective and then check to see if that language contains ambiguity.\par

@@

We are interested in how each speaker-model uses ambiguity over the discourse. A speaker strategy that is mutually efficient for both agents should avoid ambiguity until sufficient contextual information has accumulated. We should expect this to be reflected in our \textit{Full pragmatics} model who reasons about the listener and discourse history. By contrast, a speaker-optimal model who does not consider the listener should greedily use ambiguous utterances (\textit{No pragmatics} model), while a listener-optimal model should avoid ambiguity entirely (\textit{Partial pragmatics} model).\par

@@

Figure 3, panel (A) shows the empirical probability a speaker uses an ambiguous utterance as a function of discourse position. The \textit{No pragmatics} baseline uses ambiguous utterances frequently and at a constant rate over the discourse. The \textit{Partial pragmatics} baseline avoids ambiguous utterances entirely. But, the \textit{Full pragmatics} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds. Panel (C) tracks our CE objective for each model over the discourse. Note that the objective decreases for all three models, primarily driven by the listeners updating their beliefs about the actual topic ($P(C=c_{current}|D)$). However, the objective declines more quickly under the \textit{Full} and \textit{Partial pragmatics} speakers as listener agents are better able to infer the correct context. Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatics} models at the end of the discourse is driven by the reduction in speaker costs. Panel (B) tracks speaker effort, which remains constant in both \textit{No pragmatics} and \textit{Partial pragmatics} baselines. But, effort declines in the \textit{Full pragmatics} model as speakers increasingly rely on ambiguous material later in the discourse.\par

### 2019-02-02 paper/paper.tex
BP final read through, now submitting.

The approaches above typically posit efficiency measures that are
motivated by information-theoretic principles, but they typically do not
ground out in language use by interacting agents. In this work, we
derive an novel objective function from first principles of
\textit{rational language use} and show how optimizing this objective
can lead to communicatively efficient systems. We also demonstrate that
assumptions about interlocutors impact whether language properties are
used efficiently. In this way, we integrate questions of language design
and language use in a single framework.\par

@@

The Rational Speech Act framework (RSA) is a computational-level theory
of pragmatic language use, which has produced good fit to human
communication behavior across a range of language phenomena (Frank \&
Goodman, 2012; Goodman \& Frank, 2016). RSA is a formalization of
essential Gricean pragmatic principles -- agents reason about one
another and their shared context (Grice, 1975). We adopt RSA as our
representational framework to model Gricean (rational and pragmatic)
speaker-listeners in the reference game setting (see
\href{https://bit.ly/2RBSGcU}{SI}).\par

@@

Piantadosi et al. (2011) argue just the opposite, claiming that
ambiguity is an \emph{efficient} property of any communication system in
which \emph{communication is contextualized}. Simply put, it is useful
to have a language that re-uses low-cost material (has ambiguity) so
long as the cost of disambiguating the material is low. In particular,
context (or common ground) can provide useful information for
disambiguation.\par

As an example, say we have two objects (\(m_1\) and \(m_2\)), two
utterances (\(u_1\) and \(u_2\)), differing in cost, and two languages
(\(\ell_1\), \(\ell_2\)), describing different utterance-meaning
mappings. In language \(\ell_1\), the low-cost \(u_1\) can be used to
refer to both \(m_1\) and \(m_2\)
(\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\)), but the high-cost \(u_2\) cannot
be used at all (\([[u_2]]_{\ell_1} = \emptyset\)). By contrast, in
language \(\ell_2\), \(u_1\) can only refer to \(m_1\) and \(u_2\) can
only refer to \(m_2\) (\([[u_1]]_{\ell_2} = \{m_1\}\) and
\([[u_2]]_{\ell_2} = \{m_2 \}\)). While it is cheaper for a speaker to
use \(\ell_1\) (because speaking is always low-cost), it is more
difficult for a listener (because \(u_1\) is ambiguous). Crucially,
however, if context is disambiguating then the speaker can use \(u_1\)
to refer to either \(m_1\) or \(m_2\) and \(\ell_1\) should be preferred
to \(\ell_2\).\par

@@

We conduct \(N=2000\) simulations. For each simulation we enumerate the
set of \emph{valid} languages in which \(|U|=|M|=4\) (\(U\) is our set
of utterances and \(M\) our set of meanings). Recall that languages are
boolean matrices and a language \(\ell \in L\) is ``valid'' so long as
each possible meaning in \(m \in M\) can be referred to by at least one
form \(u \in U\) (every column of \(\ell\) has some non-zero assignment)
and each form maps to at least one meaning (every row has some non-zero
assignment). For a given simulation, the goal is to find the language
\(\ell^*\) which minimizes our objective and then check to see if that
language contains ambiguity.\par

@@

Figure 3, panel (A) shows the empirical probability a speaker uses an
ambiguous utterance as a function of discourse position. The
\textit{No pragmatic} baseline uses ambiguous utterances frequently and
at a constant rate over the discourse. The \textit{Partial pragmatic}
baseline avoids ambiguous utterances entirely. But, the
\textit{Full pragmatic} model avoids ambiguous material only at the
start of the discourse, employing it increasingly as the discourse
proceeds. Panel (C) tracks our CE objective for each model over the
discourse. Note that the objective decreases for all three models,
primarily driven by the listeners updating their beliefs about the
actual topic (\(P(C=c_{current}|D)\)). However, the objective declines
more quickly under the \textit{Full} and \textit{Partial pragmatic}
speakers as listener agents are better able to infer the correct
context. Additionally, the difference in CE between the \textit{Full}-
and \textit{Partial pragmatic} models at the end of the discourse is
driven by the reduction in speaker costs. Panel (B) tracks speaker
effort, which remains constant in both \textit{No pragmatic} and
\textit{Partial pragmatic} baselines. But, effort declines in the
\textit{Full pragmatic} model as speakers increasingly rely on ambiguous
material later in the discourse.\par

>>>>

The approaches above typically posit efficiency measures that are
motivated by information-theoretic principles, but they typically do not
ground out in language use by interacting agents. In this work, we
derive a novel objective function from first principles of
\textit{rational language use} and show how optimizing this objective
can lead to communicatively efficient systems. We also demonstrate that
assumptions about interlocutors impact whether language properties are
used efficiently. In this way, we integrate questions of language design
and language use in a single framework.\par

@@

The Rational Speech Act framework (RSA) is a computational-level theory
of pragmatic language use, which has produced good fit to human
communication behavior across a range of language phenomena (Frank \&
Goodman, 2012; Goodman \& Frank, 2016). RSA is a formalization of
essential Gricean pragmatic principles -- agents reason about one
another and their shared context (Grice, 1975). We adopt RSA as our
representational framework to model Gricean (rational and pragmatic)
speakers and listeners in the reference game setting (see
\href{https://bit.ly/2RBSGcU}{SI}).\par

@@

Piantadosi et al. (2011) argue just the opposite, claiming that
ambiguity is an efficient property of any communication system in which
\emph{communication is contextualized}. Simply put, it is useful to have
a language that re-uses low-cost material (has ambiguity) so long as the
cost of disambiguating the material is low. In particular, context (or
common ground) can provide useful information for disambiguation.\par

As an example, say we have two objects (\(m_1\) and \(m_2\)), two
utterances (\(u_1\) and \(u_2\)), differing in cost, and two languages
(\(\ell_1\) and \(\ell_2\)), describing different utterance-meaning
mappings. In language \(\ell_1\), the low-cost \(u_1\) can be used to
refer to both \(m_1\) and \(m_2\)
(\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\)), but the high-cost \(u_2\) cannot
be used at all (\([[u_2]]_{\ell_1} = \emptyset\)). By contrast, in
language \(\ell_2\), \(u_1\) can only refer to \(m_1\) and \(u_2\) can
only refer to \(m_2\) (\([[u_1]]_{\ell_2} = \{m_1\}\) and
\([[u_2]]_{\ell_2} = \{m_2 \}\)). While it is cheaper for a speaker to
use \(\ell_1\) (because speaking is always low-cost), it is more
difficult for a listener (because \(u_1\) is ambiguous). Crucially,
however, if context is disambiguating then the speaker can use \(u_1\)
to refer to either \(m_1\) or \(m_2\) and \(\ell_1\) should be preferred
to \(\ell_2\).\par

@@

We conduct \(N=2000\) simulations. For each simulation we enumerate the
set of \emph{valid} languages in which \(|U|=|M|=4\) (\(U\) is our set
of utterances and \(M\) our set of meanings). Recall that languages are
boolean matrices and a language \(\ell \in L\) is ``valid'' so long as
each possible meaning \(m \in M\) can be referred to by at least one
form \(u \in U\) (every column of \(\ell\) has some non-zero assignment)
and each form maps to at least one meaning (every row has some non-zero
assignment). For a given simulation, the goal is to find the language
\(\ell^*\) which minimizes our objective and then check to see if that
language contains ambiguity.\par

@@

Figure 3, panel (A) shows the empirical probability a speaker uses an
ambiguous utterance as a function of discourse position. The
\textit{No pragmatics} baseline uses ambiguous utterances frequently and
at a constant rate over the discourse. The \textit{Partial pragmatics}
baseline avoids ambiguous utterances entirely. But, the
\textit{Full pragmatics} model avoids ambiguous material only at the
start of the discourse, employing it increasingly as the discourse
proceeds. Panel (C) tracks our CE objective for each model over the
discourse. Note that the objective decreases for all three models,
primarily driven by the listeners updating their beliefs about the
actual topic (\(P(C=c_{current}|D)\)). However, the objective declines
more quickly under the \textit{Full} and \textit{Partial pragmatics}
speakers as listener agents are better able to infer the correct
context. Additionally, the difference in CE between the \textit{Full}-
and \textit{Partial pragmatics} models at the end of the discourse is
driven by the reduction in speaker costs. Panel (B) tracks speaker
effort, which remains constant in both \textit{No pragmatics} and
\textit{Partial pragmatics} baselines. But, effort declines in the
\textit{Full pragmatics} model as speakers increasingly rely on
ambiguous material later in the discourse.\par
